```{r, echo=FALSE, warning=FALSE, message=FALSE, include=FALSE}
library(synapser)
library(knit2synapse)
library(tidyverse)
library(dplyr)
library(githubr)
library(ggplot2)
library(ggpval)
library(ggExtra)
library(ggpubr)
library(data.table)
library(patchwork)

knitr::opts_chunk$set(cache = FALSE)
synLogin()
```

## About
In this assesssment we will be comparing the two groups of reported pain/no-pain using a boxplot, which we will compare it based on gait metrics based on [pdkit features](https://pdkit.readthedocs.io/en/latest/gait.html), specificially calculating step-step duration and symmetry ratio from our Psorcast cohort. Assessment are taken based on x-axis measurements, which aligns to our Walk 30 secs activity that prompts user to put their phone inside pocket verticall (walk is going in the x-axis direction of the phone)

```{r, echo = FALSE}
MERGED_FEATURES <- "syn25832975"
```

```{r echo = FALSE}
get_feature_boxplot <- function(data, feature, metric, title, x, y){
    ggplot(data=data, 
       aes_string(
           x = metric, 
           y = feature,
           fill = metric)) +
    geom_boxplot(
        alpha =  0.5,
        width = 0.5/length(unique(data[[metric]]))) +
    geom_dotplot(binaxis='y', 
                 stackdir='center', 
                 dotsize=0.7, 
                 fill = "black", 
                 alpha = 0.5) +
    ggpubr::stat_compare_means() +
    labs(
        title = title,
        x = x, 
        y = y) +
    theme_minimal() +  
    theme(plot.subtitle = 
              element_text(family = "sans", size = 4,),
          axis.text.x = 
              element_blank())
}
```

```{r include=FALSE}
data <-  fread(synGet(MERGED_FEATURES)$path)

ppacman <-  synTableQuery("SELECT *  FROM syn25006883")$asDataFrame() %>% 
    dplyr::select(participantId, 
                  visit_num = `Visit Number`, 
                  ent_loc =`Enthesitis Label`) %>%
    dplyr::mutate(has_lower_enthesitis = ifelse(str_detect(ent_loc, "knee|achilles"), 
                                                TRUE, 
                                                FALSE),
                  has_upper_enthesitis = ifelse(str_detect(ent_loc, "arm"), 
                                                TRUE, 
                                                FALSE)) %>% 
    dplyr::mutate(participantId = tolower(participantId))

data <- data %>%
    dplyr::inner_join(ppacman, by = c("participantId", "visit_num"))
```


```{r, echo = FALSE}
cleaned.data <- data %>% 
    drop_na(c("x_step_duration_md",
              "x_step_duration_iqr",
              "x_stride_deviation_md",
              "x_stride_deviation_iqr",
              "x_cadence_md",
              "x_cadence_iqr",
              "x_symmetry_md",
              "x_symmetry_iqr")) %>%
    dplyr::mutate(
        complete_case_joint_pain_list = gs_jc_joint_list,
        complete_case_swollen_joint_list = gs_swell_joint_list) %>%
    dplyr::mutate(
        complete_case_joint_pain_list = coalesce(
            complete_case_joint_pain_list,  tjc_backup),
        complete_case_swollen_joint_list = coalesce(
            complete_case_swollen_joint_list,  sjc_backup)) %>%
    dplyr::mutate(
        has_lower_joint_pain = case_when(
            stringr::str_detect(
                complete_case_joint_pain_list, 
                "knee|hip|ankle") ~ TRUE, 
            TRUE ~ FALSE),
        knee_pain = case_when(
            stringr::str_detect(
                complete_case_joint_pain_list, "knee") ~ TRUE, 
            TRUE ~  FALSE),
        hip_pain = case_when(
            stringr::str_detect(
                complete_case_joint_pain_list, "hip") ~ TRUE, 
            TRUE ~  FALSE),
        ankle_pain = case_when(
            stringr::str_detect(
                complete_case_joint_pain_list, "ankle") ~ TRUE, 
            TRUE ~ FALSE)) %>%
    dplyr::mutate(combined_lower_pain = case_when(
        (has_lower_joint_pain == TRUE | has_lower_enthesitis == TRUE) ~ TRUE,
        TRUE ~ FALSE
    )) %>%
    dplyr::select(
        participantId, age, sex, diagnosis, 
        matches("pain|step|stride|symmetry|cadence|enthesitis"))
```

### Samples of Reported Pain/No-Pain
```{r, echo = FALSE, warning = FALSE, message = FALSE}
pain_no_pain <- cleaned.data %>% 
                 dplyr::group_by(combined_lower_pain) %>% 
                 dplyr::summarise(n = n()) %>%
                 dplyr::select(status = combined_lower_pain, everything())

gender_break_down <- cleaned.data %>% 
                 dplyr::group_by(sex, combined_lower_pain) %>% 
                 dplyr::summarise(n = n()) %>%
                 dplyr::select(sex, status = combined_lower_pain, everything())

knitr::kable(list(pain_no_pain, gender_break_down))
```

### Demographics Distribution

### a. Pain vs no Pain
```{r, echo = FALSE, fig.width = 10, fig.height = 5, warning = FALSE, message = FALSE}
ggplot(data=cleaned.data %>% dplyr::filter(sex != ""), 
       aes(x=combined_lower_pain, y = age, fill = sex)) +
    geom_boxplot(width = 0.5, alpha = 0.5) +
    geom_jitter(position=position_dodge(0.5)) +
    labs(title = "Demographics Distribution (Differentiated by Gender)") +
    labs(subtitle = "Boxplot of Age vs Lower Joint Pains",
         colour = "gray") +
    theme_minimal() +  
    theme(plot.subtitle = element_text(family = "sans", size = 15),
          plot.title = element_text(family = "sans", size = 20))
```


```{r echo = FALSE}
# get features mapping
features_mapping <- list(
    x_cadence_md = list(title = "Median Cadence (Steps/Seconds)",
                        x = "",
                        y = "Steps/Seconds"),
    x_cadence_iqr = list(title = "IQR Cadence (Steps/Seconds)",
                    x = "",
                    y = "Steps/Seconds"),
    x_step_duration_md = list(title = "Median Step Duration",
                    x = "",
                    y = "Seconds"),
    x_step_duration_iqr = list(title = "IQR Step Duratioin",
                    x = "",
                    y = "Seconds"),
    x_step_deviation_md = list(title = "Median Step Deviation",
                    x = "",
                    y = "Seconds"),
    x_step_deviation_iqr = list(title = "IQR Step Deviation",
                    x = "",
                    y = "Seconds"),
    x_stride_duration_md = list(title = "Median Stride Duration",
                    x = "",
                    y = "Seconds"),
    x_stride_duration_iqr = list(title = "IQR Stride Duratioin",
                    x = "",
                    y = "Seconds"),
    x_stride_deviation_md = list(title = "Median Stride Deviation",
                    x = "",
                    y = "Seconds"),
    x_stride_deviation_iqr = list(title = "IQR Stride Deviation",
                    x = "",
                    y = "Seconds"),
    x_symmetry_md = list(title = "Median Walk Symmetry",
                    x = "",
                    y = "% Symmetric"),
    x_symmetry_iqr = list(title = "IQR Walk Symmetry",
                    x = "",
                    y = "% Symmetric")
    
)
```


### b. Walk 30secs Feature Comparison for Lower Joints Pain/No-Pain
```{r echo = FALSE, fig.width = 13, fig.height = 18, warning = FALSE, message = FALSE}
purrr::map(names(features_mapping), function(feature){
    feature_info <- features_mapping[[feature]]
    get_feature_boxplot(
        data = cleaned.data,
        feature = feature,
        metric = "has_lower_joint_pain",
        title = feature_info$title,
        x = feature_info$x,
        y = feature_info$y)}) %>% 
    patchwork::wrap_plots(ncol = 2, 
                          guides = "collect",
                          heights = 5)
```


### c. Walk 30secs Feature Comparison for Lower Joints Enthesitis Labels
```{r echo = FALSE, fig.width = 13, fig.height = 18, warning = FALSE, message = FALSE}
purrr::map(names(features_mapping), function(feature){
    feature_info <- features_mapping[[feature]]
    get_feature_boxplot(
        data = cleaned.data,
        feature = feature,
        metric = "has_lower_enthesitis",
        title = feature_info$title,
        x = feature_info$x,
        y = feature_info$y)}) %>% 
    patchwork::wrap_plots(ncol = 2, 
                          guides = "collect",
                          heights = 5)
```

### d. Walk 30secs Feature Comparison for Lower Joints Combined (Enthesitis + Lower Pain) Labels
```{r echo = FALSE, fig.width = 13, fig.height = 18, warning = FALSE, message = FALSE}
purrr::map(names(features_mapping), function(feature){
    feature_info <- features_mapping[[feature]]
    get_feature_boxplot(
        data = cleaned.data,
        feature = feature,
        metric = "combined_lower_pain",
        title = feature_info$title,
        x = feature_info$x,
        y = feature_info$y)}) %>% 
    patchwork::wrap_plots(ncol = 2, 
                          guides = "collect",
                          heights = 5)
```

