---
title: Psoriasis Draw BSA Analysis
maintainer: aryton.tediarjo@sagebase.org
createdOn: `r format(Sys.time())`
---

```{r, echo=FALSE, warning=FALSE, message=FALSE, include=FALSE}
library(synapser)
library(knit2synapse)
library(tidyverse)
library(dplyr)
library(githubr)
library(ggpval)
library(ggExtra)
library(ggpubr)
library(data.table)
library(DescTools)

knitr::opts_chunk$set(cache = FALSE)
synLogin()
```

## About
This markdown is based on Psorcast Draw Analysis where we try to assess the comparison of our digital assessment to the gold standard (clinical) based on the reported psoriasis draw area. Report will be assessing based of correlation (Spearman and Lin's CCC) and Bland-altman. 

```{r, echo = FALSE}
MERGED_FEATURES <- "syn25832975"
```

```{r, echo = FALSE}
data <-  fread(synGet(MERGED_FEATURES)$path) 
cleaned.data <- data %>% 
        dplyr::mutate(dig_bsa = dig_bsa * 100) %>%
        drop_na(c(dig_bsa, gs_bsa))
```

### Diagnosis Report
```{r, echo = FALSE, warning = FALSE, message = FALSE}
diagnosis_breakdown <- cleaned.data %>% 
                 dplyr::group_by(diagnosis) %>% 
                 dplyr::summarise(n = n()) %>%
                 dplyr::select(status = diagnosis, everything())

gender_break_down <- cleaned.data %>% 
                 dplyr::group_by(sex, diagnosis) %>% 
                 dplyr::summarise(n = n()) %>%
                 dplyr::select(sex, status = diagnosis, everything())

knitr::kable(list(diagnosis_breakdown, gender_break_down))
```

### Demographics for Analysis
```{r, echo = FALSE, fig.width = 10, fig.height = 5, warning = FALSE, message = FALSE}
ggplot(data=cleaned.data, 
       aes(x=diagnosis, y = age, fill = sex)) +
    geom_boxplot(width = 0.5, alpha = 0.5) +
    geom_jitter(position=position_dodge(0.5)) + 
    scale_x_discrete(limits=c("Control", "PsO", "PsA")) +
    labs(title = "Demographics Distribution") +
    labs(subtitle = "Boxplot of Age vs Diagnosis (Differentiated by Gender)",
         colour = "gray") +
    theme_minimal() +
    theme(plot.subtitle = element_text(family = "sans", size = 15),
          plot.title = element_text(family = "sans", size = 20))
```

### Comparing Clinical BSA (%) 
```{r, echo = FALSE, fig.width = 10, fig.height = 5, warning = FALSE, message = FALSE}
digital <- ggplot(cleaned.data %>% dplyr::filter(diagnosis %in% c('PsO', 'PsA')), 
       aes(x = dig_bsa, fill = diagnosis)) + 
    geom_density(alpha = 0.3) +
    theme_minimal() +
    labs(title = "Dig. BSA") +
    xlim(-10,100) +
    theme(plot.title = element_text(size = 15, face = "bold"))

gs <- ggplot(cleaned.data %>% dplyr::filter(diagnosis %in% c('PsO', 'PsA')), 
       aes(x = gs_bsa, fill = diagnosis)) + 
    geom_density(alpha = 0.3) +
    theme_minimal() +
    labs(title = "Gold Standard BSA") +
    xlim(-10,100) +
    theme(plot.title = element_text(size = 15, face = "bold"))

patchwork::wrap_plots(digital, gs, guides = 'collect')
```


```{r, echo = FALSE, warning = FALSE, message = FALSE}
# get correlation values
cc.corr <- DescTools::CCC(
    cleaned.data$dig_bsa, 
    cleaned.data$gs_bsa, 
    ci = "z-transform", conf.level = 0.95, na.rm = FALSE) %>% 
    .$rho.c %>% .$est

sp.corr <- cor.test(cleaned.data$dig_bsa, 
         cleaned.data$gs_bsa, 
         method = "spearman")

label <- glue::glue("Spearman = {spearman_corr}, P = {spearman_p}\n",
                    "Lin's CCC = {ccc}",
                    spearman_corr = round(sp.corr$estimate[[1]], 2),
                    spearman_p = formatC(sp.corr$p.value, digit=2, format = "e"),
                    ccc = round(cc.corr, 2))
```

### Comparing Digital vs Gold Standard Measures
```{r, echo = FALSE, fig.width = 15, fig.height = 6, warning = FALSE, message = FALSE}
#' joint pain
joint.plot.bsa <- (ggscatter(
    cleaned.data, 
    x = "dig_bsa", 
    y = "gs_bsa", 
    add = "reg.line", 
    size = 2,
    add.params = list(color = "skyblue3", fill = "lightgray"),
    conf.int = TRUE, 
    fullrange = T) + 
        scale_x_continuous(name = "Digital BSA (%)") +
        annotate("text", label = label, x = 10, y = 60, size = 4) +
        scale_y_continuous(name = "Gold Standard BSA (%)") + 
        labs(title = "Psoriasis Draw BSA (%) Correlation Plot") +
        theme_minimal() +
        theme(plot.title = element_text(
            family = "sans", vjust = -1,
            size = 20, margin=margin(0,0,10,0)))) %>%
    ggExtra::ggMarginal(., type = "densigram", 
                        xparams = list(fill = "gray"), 
                        yparams = list(fill = "gray"))

bland.altman.bsa <- blandr::blandr.draw(
    cleaned.data$dig_bsa, 
    cleaned.data$gs_bsa, 
    ciDisplay = FALSE,
    point_size = 3) + 
    scale_y_continuous(name = "Difference", expand=c(0,0), limits=c(-20,50)) +
    labs(title = "Prosiasis Draw BSA (%) Bland-Altman Plot") + 
    theme_minimal() + 
    theme(plot.title = element_text(family = "sans", vjust = -1, 
                                    size = 20, margin=margin(0,0,60,0)))

patchwork::wrap_plots(joint.plot.bsa, bland.altman.bsa)
```
