---
title: Jar Open EDA - Extended
maintainer: aryton.tediarjo@sagebase.org, meghasyam@sagebase.org
createdOn: `r format(Sys.time())`
---

## About
This notebook will explore features used in Psorcast Digital Jar Opener
as a function of joint pain and reported upper body pain from the cohort. 

- Comparing features based on uppper body pain (binary yes/no)
- Comparing features based on number of reported joint pain

```{r, echo=FALSE, warning=FALSE, message=FALSE, include=FALSE}
library(synapser)
library(knit2synapse)
library(tidyverse)
library(dplyr)
library(githubr)
library(ggplot2)
library(ggpval)
library(ggExtra)
library(ggpubr)
library(data.table)
library(patchwork)

knitr::opts_chunk$set(cache = FALSE)
synLogin()
```

```{r, echo = FALSE}
MERGED_FEATURES <- "syn25832975"
```

```{r echo = FALSE}
data <-  fread(synGet(MERGED_FEATURES)$path)
```

```{r echo = FALSE}
get_total_rotation <- function(data){
    data %>% 
        dplyr::rowwise() %>% 
        dplyr::mutate(total_rotation = sum(
            djo_leftCounter, djo_rightCounter,
            djo_leftClockwise, djo_rightClockwise)) 
}

count_joint <- function(data, group){
    output_variable <- glue::glue("{group}_status_counts")
    data %>% 
        dplyr::mutate(!!sym(output_variable) := 
                          case_when(
                              is.na(!!sym(group)) ~ 0,
                              !!sym(group)== "both" ~ 2,
                              TRUE ~ 1))
}

get_upper_body_pain <- function(data){
    data %>%
        dplyr::rowwise() %>% 
        dplyr::mutate(dig_upper_body_pain = 
                          !(is.na(dig_jc_status_wrist) & 
                                is.na(dig_jc_status_elbow) & 
                                is.na(dig_jc_status_shoulder))) %>%
        dplyr::mutate(gs_upper_body_pain = 
                          !(is.na(gs_jc_status_wrist) & 
                                is.na(gs_jc_status_elbow) & 
                                is.na(gs_jc_status_shoulder))) %>%
        dplyr::mutate(gs_upper_body_swell = 
                          !(is.na(gs_swell_status_wrist) & 
                                is.na(gs_swell_status_elbow) & 
                                is.na(gs_swell_status_shoulder)))
}

plot_boxplot <- function(data, group, value){
    data %>%
        dplyr::select(total_rotation, 
                      any_of(group),
                      any_of(value)) %>%
        ggplot(aes_string(x = group, y = value)) +
        geom_boxplot() +
        geom_jitter(position=position_jitter(0.1), alpha = 0.5) +
        ggpubr::stat_compare_means() +
        theme_minimal()
}

plot_subgroup <- function(data, group, value){
    group_comp <- tidyr::crossing(
        var1 = 0, var2 = 0:max(data[["total_reported"]])) %>%
        dplyr::slice(2:n()) %>%
        dplyr::mutate_all(as.character) %>%
        purrr::transpose(.) %>% 
        purrr::map(., function(x){unname(unlist(x))})
    data %>%
        dplyr::mutate(!!sym(group) := factor(
            !!sym(group),
            levels = c(0:max(.[[group]])))) %>%
        ggplot(aes_string(x = group, y = value)) +
        geom_boxplot() +
        geom_jitter(position=position_jitter(0.1), alpha = 0.5) +
        ggpubr::stat_compare_means(comparisons = group_comp) +
        ggpubr::stat_compare_means(label.y = 2300, label.x = "1") +
        theme_minimal()
}
```


```{r, echo=FALSE}
### Calculate required values for analysis
features <- data %>% 
    get_upper_body_pain() %>% 
    get_total_rotation()
```


### Rotation Ratios
```{r, echo = FALSE, fig.width = 8, fig.height = 10, warning = FALSE, message = FALSE}
# digital measures
dig_measures_pain <- purrr::map(c("djo_inward_ratio", "djo_outward_ratio"),
                           function(metric){
                               features %>% 
                                   plot_boxplot(
                                       group = "dig_upper_body_pain", 
                                       value = metric)}) %>%
    patchwork::wrap_plots(ncols = 2)

# gold standard
gs_measures_pain <- purrr::map(c("djo_inward_ratio", "djo_outward_ratio"),
                           function(metric){
                               features %>% 
                                   plot_boxplot(
                                       group = "gs_upper_body_pain", 
                                       value = metric)}) %>%
    patchwork::wrap_plots(ncols = 2)

# gold standard
gs_measures_swell <- purrr::map(c("djo_inward_ratio", "djo_outward_ratio"),
                           function(metric){
                               features %>% 
                                   plot_boxplot(
                                       group = "gs_upper_body_swell", 
                                       value = metric)}) %>%
    patchwork::wrap_plots(ncols = 2)

ratio_plot <- patchwork::wrap_plots(dig_measures_pain, 
                      gs_measures_pain, 
                      gs_measures_swell,
                      nrow = 3)

ratio_plot
```


### Total Rotation
```{r, echo = FALSE, fig.width = 8, fig.height = 4, warning = FALSE, message = FALSE}
# digital measures
dig_measures_pain <- purrr::map(c("total_rotation"),
                           function(metric){
                               features %>% 
                                   plot_boxplot(
                                       group = "dig_upper_body_pain", 
                                       value = metric)}) %>%
    patchwork::wrap_plots(ncols = 1)

# gold standard
gs_measures_pain <- purrr::map(c("total_rotation"),
                           function(metric){
                               features %>% 
                                   plot_boxplot(
                                       group = "gs_upper_body_pain", 
                                       value = metric)}) %>%
    patchwork::wrap_plots(ncols = 1)

# gold standard
gs_measures_swell <- purrr::map(c("total_rotation"),
                           function(metric){
                               features %>% 
                                   plot_boxplot(
                                       group = "gs_upper_body_swell", 
                                       value = metric)}) %>%
    patchwork::wrap_plots(ncols = 1)

rotation_plot <- patchwork::wrap_plots(dig_measures_pain, 
                      gs_measures_pain, 
                      gs_measures_swell,
                      ncol = 3)

rotation_plot
```

### Dig Joint Pain  - Reported Upper Body Pain to Total Rotation
```{r, echo = FALSE, fig.width = 8, fig.height = 5, warning = FALSE, message = FALSE}
metrics <- c("dig_jc_status_wrist",
             "dig_jc_status_elbow",
             "dig_jc_status_shoulder")
dig_pain <- purrr::map(metrics, ~features %>% 
               count_joint(group = .x)) %>% 
    purrr::reduce(full_join) %>% 
    dplyr::rowwise() %>%
    dplyr::mutate(total_reported = sum(across(ends_with("status_counts")))) %>% 
    dplyr::select(total_reported, total_rotation)
dig_pain %>%
    plot_subgroup(group = "total_reported", 
                  value = "total_rotation") + 
    labs(x = "Total Reported Dig. Upper-Body Pain")
```


### Gold Standard Joint Pain - Reported Upper Body Pain to Total Rotation
```{r, echo = FALSE, fig.width = 8, fig.height = 5, warning = FALSE, message = FALSE}
metrics <- c("gs_jc_status_wrist",
             "gs_jc_status_elbow",
             "gs_jc_status_shoulder")
gs_pain <- purrr::map(metrics, ~features %>% 
               count_joint(group = .x)) %>% 
    purrr::reduce(full_join) %>% 
    dplyr::rowwise() %>%
    dplyr::mutate(total_reported = sum(across(ends_with("status_counts")))) %>% 
    dplyr::select(total_reported, total_rotation)
gs_pain %>%
    plot_subgroup(group = "total_reported", 
                  value = "total_rotation") + 
    labs(x = "Total Reported Gold-Standard Upper-Body Pain")
```


### Gold Standard Swelling - Reported Upper Body Pain to Total Rotation
```{r, echo = FALSE, fig.width = 8, fig.height = 5, warning = FALSE, message = FALSE}
metrics <- c("gs_swell_status_wrist",
             "gs_swell_status_elbow",
             "gs_swell_status_shoulder")
gs_swell <- purrr::map(metrics, ~features %>% 
               count_joint(group = .x)) %>% 
    purrr::reduce(full_join) %>% 
    dplyr::rowwise() %>%
    dplyr::mutate(total_reported = sum(across(ends_with("status_counts")))) %>% 
    dplyr::select(total_reported, total_rotation)
gs_swell %>%
    plot_subgroup(group = "total_reported", 
                  value = "total_rotation") + 
    labs(x = "Total Reported Gold-Standard Upper-Body Swelling")
```